---
title: "Zubenelgenubi's Performance during BYC Mack 2018"
output: html_notebook
---

```{r include=FALSE}
rm(list=ls())

library(ggplot2)
library(ggrepel)
library(tidyverse)
library(lubridate)

#load data
load("../data/marks.Rdata")           #location of the marks
load("../data/zube_performance.rda")  #zube's performance variables
load("../data/results.Rdata")         #class data from BYCMACK webpage
load("../data/positions_wcorr.Rdata") #yellowbrick positions with corrected time
load("../data/zube_yb_withdiff.rda")  #zube yellowbrick data with difference from class0 boats


#Data Cleanup and Joining

#add class data to YellowBrick position data
results <- results %>% select(-elapsed_time, -Corrected)
  
#extract our class
classo <- posn %>% filter(class == "Class O Class")



```

## Objectives
* Where did we gain and lose versus other boats
* Where we sailing the boat optimally 

## Key Performance Variables

###Output
* Gain/loss from other boats

###Inputs
* Boatspeed (BSP) - Are we hitting our target boatspeed
* Course (COG) - Are we navigating the optimal course to get us closest to the mark. 
* Velocity Made Course (VMC) - Velocity towards the mark


## Course Sailed
Here is the full course sailed by Zube. 

```{r}

#get states shape files
mi <- map_data("state") %>% filter(region=="michigan")


cols <- c("Class O"="grey", "Expedition Data"="blue", "YellowBrick Data"="red")

#plot the positions
ggplot()+
  geom_polygon(data=mi, aes(long, lat, group=group)) + 
  coord_map("polyconic", xlim=c(-85, -82), ylim=c(42.8, 46)) + 
  geom_path(data=classo, aes(x=lon, y=lat, color="Class O")) + 
  geom_point(data=zube, aes(x=lon.exp, y=lat.exp, color="Expedition Data"))+
  geom_path(data=zube_yb, aes(x=lon, y=lat, color="YellowBrick Data", group="boat"))+
  scale_color_manual(name="Data Source", values=cols)
  
  
  


```

###split by legs 
In order to view the data more precisely we're going to split the race into 4 legs: 

* Start to Harbor Beach
* Harbor Beach to ThunderBay Light
* ThunderBay Light to Presque Isle
* Presque Isle to Finish

```{r message=FALSE, warning=FALSE}

legs <- data.frame(leg = c("ST-HB", "ST-HB", "HB-TBLT", "TBLT-PI", "TBLT-PI", "TBLT-PI", "PI-FIN", "PI-FIN"), 
                   mark = unique(zube$mark))

legs$leg <- ordered(legs$leg, levels = c("ST-HB", "HB-TBLT", "TBLT-PI", "PI-FIN"))

zube <- left_join(zube, legs)

zube_yb <- left_join(zube_yb, legs)

posn <- left_join(posn, legs)

classo <- left_join(classo, legs)


#plot the positions
ggplot()+
  geom_polygon(data=mi, aes(long, lat, group=group)) + 
  coord_map("polyconic", xlim=c(-85, -82), ylim=c(42.8, 46)) + 
  geom_point(data=zube, aes(x=lon.exp, y=lat.exp, color="Expedition"), color="blue")+
  geom_path(data=zube_yb, aes(x=lon, y=lat, color="YellowBrick"), color="red") + 
  facet_grid(. ~ leg)

```

#Leg 1 Start to Harbor Beach

## Our Path versus Everyone Else
```{r}

classdat <- classo %>% filter(place != "RET" & leg == "ST-HB")
zubeybdat <- zube_yb %>% filter(leg == "ST-HB")
zubedat <- zube %>% filter(leg == "ST-HB")

times <- seq(min(classdat$time), max(classdat$time), by="6 hour")

hrpoints <- classdat[classdat$time %in% times, ]

ggplot()+
  geom_polygon(data=mi, aes(long, lat, group=group)) + 
  coord_map("polyconic", xlim=c(-82.7, -82.25), ylim=c(43.05, 43.9)) + 
  geom_path(data=classdat, aes(lon, lat, color=boat))+
  geom_point(data= hrpoints, aes(lon,lat, color=boat, shape=as.factor(time)))+
  geom_path(data=zubeybdat, aes(lon, lat, color="zube"), color="black")



```


Gain and Loss versus our conversation
```{r}

HMformat <- function(x){ format(x, "%R")}

xaxistime <- seq(min(classdat$time), max(classdat$time), by="2 hour")
xaxislabels <- as.character(format(xaxistime, "%a %R"))

ggplot(classdat) + 
  geom_path(aes(x = time, y=dmg, color=boat)) + 
  geom_path(data=classdat[classdat$boat=="Zubenelgenubi", ],  aes(x = time, y=dmg, color="Zube"), color="black")+
  geom_vline(xintercept=times, color="Dark Gray") + 
  scale_x_time(breaks = xaxistime, labels=xaxislabels)+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  ggtitle("Distance Made Good to the Finish")

```

Gain vs. rest of the fleet
```{r}

zubeybdat$gainloss <- c(diff(zubeybdat$vsAlbacore) ,0)/.25 * -1

hrpoints <- zubeybdat[zubeybdat$time %in% times, ]

ggplot(zubeybdat) + geom_path(aes(x=time, y=vsAlbacore*-1, color=gainloss), size=2) +
  geom_vline(xintercept=times, color="Dark Gray") + 
  #scale_color_viridis(name = "Loss vs. Albacore [kts]")+
  scale_color_gradient2(high="blue", low="red", midpoint=0, name="Gain vs. Albacore [kts]")+
  scale_x_time(breaks = xaxistime, labels=xaxislabels)+
  ggtitle("Zube vs. Albacore Gain/Loss in Corrected DMG")+
  xlab("Time")+
  ylab("Distance Made Good (NM) -- Corrected")+
  theme(axis.text.x = element_text(angle = 60, hjust = 1))+
  geom_hline(yintercept=0)

ggplot(zubeybdat) + 
  geom_polygon(data=mi, aes(long, lat, group=group)) + 
  coord_map("polyconic", xlim=c(-82.7, -82.25), ylim=c(43.05, 43.9)) + 
  geom_path(aes(x=lon, y=lat, color=gainloss), size=2)+
  geom_point(data=hrpoints, aes(x=lon, y=lat, color=gainloss))+
  geom_text(data=hrpoints, aes(x=lon, y=lat, label=format(time, "%a %R")), nudge_x = 0.1)+
  scale_color_gradient2(high="blue", low="red", midpoint=0, name = "Gain vs. Albacore [kts]")
  #scale_color_viridis(name = "Loss vs. Albacore [kts]")

```

Performance over that same range
```{r}



ggplot(zube[zube$gainrate < 2 & zube$gainrate > -2, ]) +
  geom_polygon(data=mi, aes(long, lat, group=group)) + 
  coord_map("polyconic", xlim=c(-82.7, -82.25), ylim=c(43.05, 43.9)) + 
  geom_path(aes(x=lon.exp, y=lat.exp, color=gainrate))+
  scale_color_gradient2(high="blue", low="red", midpoint=0, name = "Gain vs. Albacore [kts]")

ggplot(zube[zube$gainrate < 2 & zube$gainrate > -2, ]) + geom_point(aes(x=gainrate, y=pol_perc)) + scale_y_continuous(limits=c(0,150))

```

Performance 
```{r}

ggplot(zube) + geom_point(aes(x=time, y=pol_perc)) + scale_y_continuous(limits = c(0,200))

```


```{r}

ggplot(zube) + geom_point(aes(x=time, y=vmc_perc)) + scale_y_continuous(limits = c(0,200))


```

